\normalfont\spacedallcaps{}

### 2019-02-09 Document/article_4.tex
third

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------
\title{\normalfont\spacedallcaps{Article Title}} % The article title

\author{\spacedlowsmallcaps{Fatemeh Hadi Nezhad\textsuperscript{1}}}

\date{2019} % An optional date to appear under the author(s)

@@

\maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

@@

\let\thefootnote\relax\footnotetext{* \textit{Department of Quantitative System Biology, University of California, Merced, United States}}

@@

\section{Preparing Data}
%------------------------------------------------
\textbf{TryTryp genome data}. From tritrypdb website, we downloaded the version 41 of 46 TryTryp genomes released on 2018-12-05. Genomes are compared based on number of sequence fragments relative to their length as shown in figure ~\vref{fig:gallery}. 
All the results, scripts and input data for TryTryp version 41 can be found \href{thesentence}{here}.

\begin{figure}[tb]
\centering 
\includegraphics[width=0.8\columnwidth]{GenomeComparisonV41.png} 
\caption[Genome Comparison]{Comparing genomes based on formula ${(\frac{f(x)}{\max(f(x))})}^{-1}$ which f(x) = Number of sequences in genome devided by the length of genome. the higher the bar is the better genome is sequenced.} % The text in the square bracket is the caption for the list of figures while the text in the curly brackets is the figure caption
\label{fig:gallery} 
\end{figure}

\subsection{\textbf{tRNA gene annotation}}
\subsubsection*{tRNA gene prediction}
In order to annotate the tRNA genes for the sequenced TryTryp genomes, we used two computational methods for tRNA prediction, tRNAscan-SE and Aragorn. We integrated the result of both genefinders by keeping the union of tRNA gene predictions generated by tRNAscan-SE v2.0 using defult options (Lowe and Eddy 1997) and Aragorn v1.2.38 using options -i116 -t -br -seq -w -e -l -d (Laslett and Canback 2004). Genes with overlapped coordinate were considered one gene. However, the identity and exact coordinate of both genefinders we saved seperately to be compared later.

@@

\subsubsection*{Initiator tRNA prediction}
We predicted the initiator tRNAs for the genes with anticodon 'CAT' from intersection of both tRNAscan (TSE) and Aragorn (ARA) Based on Conserved positions of initiators in Eukarya from the study by CHRISTIAN MARCK and HENRI GROSJEAN. Based on the fallowing citeria:
\begin{enumerate}[noitemsep] % [noitemsep] removes whitespace between the items for a compact look
\item In all eukaryotic tDNA-iMet, positions 11–24 are occupied by C-G, However, eukaryotic elongators
also prefer C-G at these positions.

@@

\item At position 20, A is strictly conserved in all eukaryotic initiators
\end{enumerate} 
To investigate all these features we clustered CAT tRNA genes using Levenshtein (edit) distance between gene sequences and Ward.D2 method to measure the dissimilarity between each two clusters. We ended up with three clusters. Table \ref{table:1} investigates each of these features in each column. from this table we see that only tRNA genes in cluster 1 have almost all the conserved features for eukaryotic initiators.

@@

To investigate and compare tRNA genes predicted by two gene finders TSE and ARA, we made four sets of genes. Set one, TSE and ARA intersection, Set two, TSE and ARA union, Set three, genes found by ARA and Set four, genes found by TSE . for intersection set, we dismissed genes which had different identity by ARA and TSE. for union set, we picked TSE identity over ARA. Table  ~\ref{table:2} shows a summary of these four sets. Further, to compare the coordinates of genes annotated by ARA and TSE we made a heatmap shown in figure ~\ref{fig:heatmap}. We see from this figure that ... 
%----------------------------------------------------------------------------------------
%	???? Analyse why we have the differences ....
%----------------------------------------------------------------------------------------

Figure ~\ref{fig:counts} shows the number of genes annotated for each genome, and Figure ~\ref{fig:types} shows which tRNA functional classes were not annotated by both TSE and ARA for each genome.

\begin{table}[hbt]
\caption{summary of the predicted genes by TSE and ARA. pseudo genes are marked as \$, initiators as X, stop as \#, sup as "?", sec as Z and pyl as O}
\begin{adjustbox}{width=\columnwidth,center}
%\caption{Table of Grades}
%\centering
% make it as two table
\begin{tabular}{|l|lllllllllllllllllllllllllllllllllll|}
\hline
GeneSet & \# tRNA & \# nucleotides & N/T & gene length & \%G & \%C & \%T & \%A & \%intron & A & C & D & E & F & G & H & I & K & L & M & N & P & Q & R & S & T & V & W & Y & X & Z & \$ & ? & \# & O\\
TSE2 & 3629 & 270211 & 74.46 & 50-164 & 31.99 & 26.11 & 23.22 & 18.68 &  2.618 & 214 & 64 & 104 & 162 & 110 & 234 &  80 & 179 & 190 & 338 & 108 & 126 & 201 & 162 & 350 & 238 & 219 & 241 & 52 &  94 & 76 & 78 & 28 & 3 & 0 & 0\\
ARA & 4345 & 372392 & 85.71 & 70-215 & 32.64 & 26.92 & 22.87 & 17.57 & 14.684 & 257 & 86 & 123 & 192 & 125 & 339 & 129 & 213 & 194 & 393 & 101 & 153 & 228 & 175 & 420 & 362 & 248 & 282 & 60 &  90 & 76 & 82 &  0 & 0 & 2 & 4\\
UNION & 4379 & 377587 & 86.23 & 50-215 & 32.81 & 26.66 & 22.87 & 17.65 & 15.346 & 259 & 86 & 118 & 193 & 130 & 344 & 129 & 220 & 197 & 380 & 112 & 143 & 229 & 175 & 421 & 369 & 249 & 282 & 57 & 106 & 76 & 82 & 28 & 3 & 2 & 2\\
INTERSECTION & 3560 & 265016 & 74.44 & 68-89 & 32.01 & 26.13 & 23.22 & 18.64 &  2.331 & 212 & 64 & 104 & 161 & 105 & 229 &  80 & 172 & 187 & 338 &  97 & 125 & 200 & 162 & 349 & 230 & 218 & 241 & 52 &  78 & 76 & 78 &  6 & 0 & 0 & 0\\
\hline
\end{tabular}
\label{table:2}
\end{adjustbox}
\end{table}

@@

\begin{figure}[tb]
\centering 
\includegraphics[width=0.8\columnwidth]{ara_funcPerc.png} 
\caption[Genome Comparison]{Percentage of 22 tRNA types annotated by both TSE and ARA for each TryTryp genomes. The labels on top of each bar shows which tRNA classes are not annotated for the genome.} % The text in the square bracket is the caption for the list of figures while the text in the curly brackets is the figure caption
\label{fig:types} 
\end{figure}

....
%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------
\newpage
\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading

>>>>

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------
\title{\normalfont\spacedallcaps{}} % The article title

%\author{\spacedlowsmallcaps{Fatemeh Hadi Nezhad\textsuperscript{1}}}

%\date{2019} % An optional date to appear under the author(s)

@@

%\maketitle % Print the title/author/date block

\setcounter{tocdepth}{3} % Set the depth of the table of contents to show sections and subsections only

@@

%\let\thefootnote\relax\footnotetext{* \textit{Department of Quantitative System Biology, University of California, Merced, United States}}

@@

\section{Preparing Data}
%------------------------------------------------
\textbf{TryTryp genome data}. From tritrypdb website, we downloaded the version 41 of 46 TryTryp genomes released on 2018-12-05. Genomes are compared based on number of sequence fragments relative to their length as shown in figure ~\vref{fig:gallery}. 
All the results, scripts for TryTryp version 41 can be found \href{https://github.com/fhadinezhadUC/Leishmania_2019}{here}.

\begin{figure}[tb]
\centering 
\includegraphics[width=0.8\columnwidth]{GenomeComparisonV41.png} 
\caption[Genome Comparison]{Comparing genomes based on formula ${(\frac{f(x)}{\max(f(x))})}^{-1}$. f(x) = Number of sequences in genome devided by the length of genome. The length of the bars shows how good the genomes are sequenced.} % The text in the square bracket is the caption for the list of figures while the text in the curly brackets is the figure caption
\label{fig:gallery} 
\end{figure}

\subsection{\textbf{tRNA gene annotation}}
\subsubsection{tRNA gene prediction}
In order to annotate tRNA genes for the sequenced TryTryp genomes, we used two computational methods for tRNA prediction, tRNAscan-SE and Aragorn. We integrated the result of both genefinders by keeping the union of tRNA gene predictions generated by tRNAscan-SE v2.0 using defult options (Lowe and Eddy 1997) and Aragorn v1.2.38 using options -i116 -t -br -seq -w -e -l -d (Laslett and Canback 2004). Genes with overlapped coordinate were considered as one gene. However, the identity and exact coordinate of both genefinders we saved seperately to be analysed later.

@@

\subsubsection{Initiator tRNA prediction}
We predicted the initiator tRNAs for the genes with anticodon 'CAT' from intersection of both tRNAscan (TSE) and Aragorn (ARA) Based on Conserved positions of initiators in Eukarya from the study by CHRISTIAN MARCK and HENRI GROSJEAN. Based on this study we have the fallowing citeria for initiators:
\begin{enumerate}[noitemsep] % [noitemsep] removes whitespace between the items for a compact look
\item In all eukaryotic tDNA-iMet, positions 11–24 are occupied by C-G, However, eukaryotic elongators
also prefer C-G at these positions.

@@

\item At position 20, A is strictly conserved in all eukaryotic initiators
\end{enumerate} 
To investigate all these features we clustered CAT tRNA genes using Levenshtein (edit) distance between gene sequences and Ward.D2 method to measure the dissimilarity between each two clusters. We ended up with three clusters. Table \ref{table:1} investigates each of these features in each column. from this table we see that only tRNA genes in cluster 1 have almost all the conserved features for eukaryotic initiators. So, we marked these genes as initiators represented with letter X in our gene file.

@@

To investigate and compare tRNA genes predicted by two gene finders TSE and ARA, we made four sets of genes. Set one, TSE and ARA intersection, Set two, TSE and ARA union, Set three, genes found by ARA and Set four, genes found by TSE . for intersection set, we dismissed genes which had different identity by ARA and TSE. for union set, we picked TSE identity over ARA. Table  ~\ref{table:2} shows a summary of these four sets. Further, to compare the coordinates of genes annotated by ARA and TSE we made a heatmap shown in figure ~\ref{fig:heatmap}. We see from this figure that the coordinates of same genes annotated by ARA and TSE do not always match. We Analysed the reason for each set of displacementin this figure as fallow:

\begin{enumerate}[noitemsep]

\item Genes with 0 displacement in both ends: These genes have same identity for both TSE and ARA except for 33 genes with Anticodon loop of more than 8bp. Also, both ARA and TSE reported genes up to base 73. 
\item Genes with 0 displacement at 5 prime end and 1 displacement at 3 prime end: Identity of these genes matches between ARA and TSE. They all have anticodon loop of length 7. The reason for the displacement in 3 prime end is that ARA reports up to position 74, however, TSE reports only up to position 73.  
\item Genes with one base displacement at both ends: In this case ARA reports one extra base at both ends which is because these two bases pair together and in most cases this makes the Amino Acid arm one base longer than what TSE reports. Although in a few cases, the AminoAcid arm will stay 7bp, but we see insertions in this arm.  
\item Genes with two base displacement at 3 prime end and one at 5 prime end: In this case ARA is reporting 2 extra bases at one end and 1 for the other end which usually leads to a longer AminoAcid arm. Also, the extra 2 bases reported by ARA at 3 prime end are mostly 'ac'.  
\item Genes with two base displacement at 3 prime end and 0 displacement at 5 prime end: In this case TSE reports up to position 73 as always, but ARA is reporting genes up to position 75. the last three bases of ARA are mainly 'acc'
\item Genes with three base displacement at 3 prime end and 0 displacement at 5 prime end. In this case ARA is reporting 3 extra bases at 3 prime end and these three bases are '?cc'.

\end{enumerate}

%----------------------------------------------------------------------------------------
%	???? Analyse why we have the differences ....
%----------------------------------------------------------------------------------------

We expect our genefinders to annotate tRNA genes of all 22 functional classes for each genome. To investigate this we visualized the number of genes annotated for each genome in Figure ~\ref{fig:counts} and tRNA functional classes annotated by both TSE and ARA for each genome in Figure ~\ref{fig:types}.

\begin{table}[hbt]
\caption{summary of the predicted genes by TSE and ARA. We marked pseudo genes as \$, initiators as X, stop as \#, sup as "?", sec as Z and pyl as O}
\begin{adjustbox}{width=\columnwidth,center}
%\caption{Table of Grades}
%\centering
% make it as two table
\begin{tabular}{|l|lllllllllllllllllllllllllllllllllll|}
\hline
GeneSet & \# tRNA & \# nucleotides & N/T & gene length & \%G & \%C & \%T & \%A & \%intron & A & C & D & E & F & G & H & I & K & L & M & N & P & Q & R & S & T & V & W & Y & X & Z & \$ & ? & \# & O\\
TSE2 & 3631 & 270355 & 74.46 & 50-164 & 31.99 & 26.11 & 23.22 & 18.68 &  2.616 & 214 & 64 & 105 & 163 & 110 & 234 &  80 & 179 & 190 & 338 & 108 & 126 & 201 & 162 & 350 & 238 & 219 & 241 & 52 &  94 & 76 & 78 & 28 & 3 & 0 & 0\\
ARA & 4347 & 372539 & 85.70 & 70-215 & 32.64 & 26.92 & 22.87 & 17.57 & 14.677 & 257 & 86 & 124 & 193 & 125 & 339 & 129 & 213 & 194 & 393 & 101 & 153 & 228 & 175 & 420 & 362 & 248 & 282 & 60 &  90 & 76 & 82 &  0 & 0 & 2 & 4\\
UNION & 4381 & 377734 & 86.22 & 50-215 & 32.81 & 26.66 & 22.87 & 17.65 & 15.339 & 259 & 86 & 119 & 194 & 130 & 344 & 129 & 220 & 197 & 380 & 112 & 143 & 229 & 175 & 421 & 369 & 249 & 282 & 57 & 106 & 76 & 82 & 28 & 3 & 2 & 2\\
INTERSECTION & 3562 & 265160 & 74.44 & 68-89 & 32.01 & 26.13 & 23.22 & 18.64 &  2.330 & 212 & 64 & 105 & 162 & 105 & 229 &  80 & 172 & 187 & 338 &  97 & 125 & 200 & 162 & 349 & 230 & 218 & 241 & 52 &  78 & 76 & 78 &  6 & 0 & 0 & 0\\
\hline
\end{tabular}
\label{table:2}
\end{adjustbox}
\end{table}

@@

\begin{figure}[tb]
\centering 
\includegraphics[width=0.8\columnwidth]{ara_funcPerc.png} 
\caption[Genome Comparison]{Percentage of 22 tRNA types annotated by both TSE and ARA for each TryTryp genomes. The label on top of each bar shows which tRNA classes are not annotated for the genome.} % The text in the square bracket is the caption for the list of figures while the text in the curly brackets is the figure caption
\label{fig:types} 
\end{figure}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------
\newpage
\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading
